---
title: "R Notebook"
output: html_notebook
---

# Part2 (1/2)

## draw graph from data

If you are starting from this file, it will be safe to set the local path by executing the code below.

```{r}
# directory choose
dir.choose <- function() {
  system("osascript -e 'tell app \"RStudio\" to POSIX path of (choose folder with prompt \"Choose Folder:\")' > /tmp/R_folder",
         intern = FALSE, ignore.stderr = TRUE)
  p <- system("cat /tmp/R_folder && rm -f /tmp/R_folder", intern = TRUE)
  return(ifelse(length(p), p, NA))
}

local <- dir.choose()
setwd(local)
```


# Read csv file

1. read the csv file you made from part1.
2. set column name for manipulation.

```{r}
inport.file.name = "1108_gem_verte_long.csv"
data <- read.csv(paste0(local, inport.file.name))
data
```

# delete row, column columns

It is not mandatory but "row" and "column" columns are not needed so delete.<br>
Change units(mM to uM) so that it will be compairable. 


```{r}
data[, c('row', 'column')] <- list(NULL) # delete 'row', 'column' columns
data[data$conc=="1mM","conc"] = "1000uM"
data
```

# Transform data for IC50

1. calculate median for each concentration
2. set control(0uM) value for reference
3. calculate viability from division of OD value
4. separate "conc" column to concentration value and unit for plot 
5. subgroup data by drug type

```{r}
if (!require("dplyr")) {
  install.packages("dplyr")
}
library(dplyr)
library(stringr)

# 1.
summary <- data %>%
  group_by(drug, conc) %>%
  summarize_at(vars(OD.value), list(value=mean))
summary

# 2.
gemcitabine.control <- summary[summary$drug=="Gemcitabin" & summary$conc=="0uM", "value"]
verteporfin.control <- summary[summary$drug=="Verteporfin" & summary$conc=="0uM", "value"]

summary["reference"] = 0
summary[summary$drug=="Gemcitabine", 'reference'] = gemcitabin.control
summary[summary$drug=="Verteporfin", 'reference'] = verteporfin.control

# 3.
summary["percent"] = summary["value"] / summary["reference"]

# 4.
summary <- summary %>%
  mutate(concentration = str_extract(conc, "^(\\d)(\\.)(\\d)+|^(\\d)+")) %>% # the order is important.
  mutate(unit = str_extract(conc, "(\\D)+$"))
summary$concentration = as.double(summary$concentration)

# 5.
gemcitabine.summary <- summary[summary$drug=="Gemcitabine",]
verteporfin.summary <- summary[summary$drug=="Verteporfin",]

gemcitabine.summary
verteporfin.summary
```


# Draw graph
https://envrisklab.hatenablog.com/entry/2018/02/12/154639
https://rstudio-pubs-static.s3.amazonaws.com/63117_47264132789f40f690a707ee5429f45a.html

Start from simple scatter plot to see if the plot shape is sigmoid curve

```{r}
plot(percent~concentration,gemcitabine.summary ,subset=concentration!=0,log="x")
# use log for x-axis
plot(percent~concentration,verteporfin.summary ,subset=concentration!=0,log="x")
```

# Analyze in detail

Lets write sigmoid graph.<br>

can add serveral sigmoid curve by setting add=TRUE
type: "all", "bars", "none", "obs"

Calculate IC50<br>
- https://www.researchgate.net/post/How_can_I_calculate_IC50_value_with_Rstudio

```{r}
if (!require("drc")) {
  install.packages("drc")
}
library(drc)

analyze <- function(title, data, line.color) {
  model.LL4 <- drm(formula=percent~concentration,
                   data=data,
                   fct=LL.4(names=c("hill", "min", "max", "ec_50")))
  summary(model.LL4)
  coef(model.LL4)
  modelFit(model.LL4)
  plot(model.LL4, main=title, conName="Control", broken = TRUE)
  abline(h = 0.5, col=line.color,  lty = 3)
  
  IC.50 <- ED(model.LL4,
              respLev=c(50),
              reference=c("control"),
              type="relative",
              bound=TRUE,
              od=FALSE,
              vocv. = vcov,
              display=TRUE,
              pool=TRUE,
              logBase=NULL,
              multiComp=FALSE)
  ic.50.value <- IC.50[1,1]
  abline(v = ic.50.value, col = line.color, lty = 3)
}

analyze("Gemcitabine IC50", gemcitabine.summary, "blue")
analyze("Verteporfin IC50", verteporfin.summary, "blue")
```

To export graph, use command line below and type "analyze("Gemcitabine IC50", gemcitabine.summary, "blue")"<br>
Graph would appear in right bottom window of RStudio.

