---
title: "[Pre-proccess]"
output: html_notebook
---

# Part (1/2)

## Preparation

You need to process excel file manually first

 1. calculate average blank value
 2. subtract blank from all cells
 3. for cells with out condition set value of 0
 4. export in a csv format and save in this downloaded folder

 Look into "20230116_MTT_1108_Met_Verte.xlsx" for actual process.<br>
 Check output csv file content by compairing it with "1108_Gem_Verte.csv"


# Set working directory

Normally, you need to just select this downloaded folder directory

```{r}
# directory choose
dir.choose <- function() {
  system("osascript -e 'tell app \"RStudio\" to POSIX path of (choose folder with prompt \"Choose Folder:\")' > /tmp/R_folder",
         intern = FALSE, ignore.stderr = TRUE)
  p <- system("cat /tmp/R_folder && rm -f /tmp/R_folder", intern = TRUE)
  return(ifelse(length(p), p, NA))
}

local <- dir.choose()
setwd(local)
```

# Read csv file

1. read the made csv file 
2. set column name to the loaded dataframe

```{r}
inport.file.name = "1108_Gem_Verte.csv"
data <- read.csv(paste0(local, inport.file.name))

setRowCol<- function(data) {
  colnames(data) <- c("row", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
  data$row <- factor(data$row)
  return (data)
} 

data <- setRowCol(data)
data
```


# convert wide to long

This is to make it easy to group & filter data and apply functions like mean.

```{r}
if (!require("tidyverse")) {
  install.packages("tidyverse") # if not installed include this line
}
library("tidyverse")

convert.to.long <- function(data) {
  data.long <- gather(data, key="column", value="OD value", 2, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13)
  return (data.long)
}

data.long <- convert.to.long(data)
data.long
```


# add settings data (adjust to your own environment)

For my data("1108 Gem Verte"), added concentration, drug type.

1. drug concentration
2. drug type

```{r}
add.data <- function(data) {
  
  # set data by row
  data["conc"] = ""
  row <- unique(data$row)
  for (i in 1:length(row)) {
    if (i == 1) {
      data[data$row == row[i],"conc"] = "blank" 
    } else if (i == 2) {
      data[data$row == row[i],"conc"] = "0uM" 
    } else if (i == 3) {
      data[data$row == row[i],"conc"] = "0.01uM" 
    } else if (i == 4) {
      data[data$row == row[i],"conc"] = "0.1uM" 
    } else if (i == 5) {
      data[data$row == row[i],"conc"] = "1uM" 
    } else if (i == 6) {
      data[data$row == row[i],"conc"] = "10uM" 
    } else if (i == 7) {
      data[data$row == row[i],"conc"] = "100uM" 
    } else { ## } if (i == 8) {
      data[data$row == row[i],"conc"] = "1mM" 
    }
  }
  
  # set data by column
  data["drug"] = ""
  column <- unique(data$column)
  for (i in 1:length(column)) {
    if (i <= 6) {
      data[data$column == i,"drug"] = "Gemcitabine"
    } else {
      data[data$column == i,"drug"] = "Verteporfin"
    }
  }

  return (data)
}

data.long <- add.data(data.long)
```

# Clean data

1. delete 0 value cells
2. convert negative values to 0 
(negative values(=viability) will be misleading)

```{r}
data.long <- subset(data.long, data.long["OD value"] != 0.0)
data.long[data.long['OD value'] < 0.0,'OD value'] = 0.0 
data.long
```

# write to file

change file name to whatever you prefer.

```{r}
output.file.name <- "1108_gem_verte_long.csv"
write_csv(data.long, paste0(local, output.file.name))
```

now you have completed converting wide format to long format.
Proceed to the next script "writeGraph.Rmd"
